---
title: "Bioinformaticsd Class 13"
author: "Pallas Duong"
date: "5/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##1000 Genomes Data Analysis 

Downloaded genotype data from Ensemble for Asthma associated SNP rs8067378 
See: (MXL) Mexican Ancestry in Los Angeles, California 

```{r}
#Read CSV file
genotype <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(genotype)
```

Find how many of each genotype there are in this dataset

```{r}
genotype.summary <- table(genotype[,2])
genotype.summary
```

Proportion of each genotype
```{r}
genotype.summary/nrow(genotype)*100
```

**Q5**: What proportion of the Mexican Ancestry in Los Angeles sample population (MXL)
are homozygous for the asthma associated SNP (G|G)? 

Answer: 14%

##RNA-seq

Look more closely at sample HG00109 with G|G genotype for this SNP (from GBR population)

##Side-note on Fastq quality scores 
```{r}
install.packages("seqinr")
install.packages("gtools")

library(seqinr)
library(gtools)
phred <- asc( s2c("DDDDCDEDCDDDDBBDDDCC@") ) - 33
phred 
```

```{r}
prob <- 10**(-phred/10)
prob
```

##RNA-Seq population sclae analysis

Read our results file from RNA-Seq across the ~230 sample (people)

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
```

Look at data 
```{r}
table(expr$geno)
```
How many samples?
```{r}
nrow(expr)
```

Expression level summary stats
```{r}
summary(expr)
```

```{r}
summary(expr$exp[(expr$geno == "G/G")])
summary(expr$exp[(expr$geno == "A/A")])
summary(expr$exp[(expr$geno == "A/G")])

```

```{r}
#boxplot
boxplot(exp ~ geno, data = expr, notch = TRUE)
```

```{r}
install.packages("ggplot2")
library(ggplot2)
ggplot(expr, aes(geno, exp)) + geom_boxplot()
```

